---
title: "p8105_hw6_bl2836"
author: "Ngan Le"
date: "2023-12-01"
output: github_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(rvest)
library(dplyr)
library(patchwork)
```

# Question 1

##### Import dataset from CSV file. 

```{r import homicide data, message = FALSE}
homicide = read_csv("homicide-data.csv")
```

This dataset includes `r nrow(homicide)` observations and `r ncol(homicide)` variables, reporting the date, the city, the state, the coordinates of the incidences, the victims' name, age, race, gender, and the disposition of the case. The incidences are reported across `r n_distinct(homicide$state)` states from 2007-2017. 

##### Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r create city_state & binary homicide vars}
cleaned_homicide =
  homicide %>%
  mutate(city_state = str_c(city, ", ", state)) %>% 
  mutate(solved = ifelse(disposition %in% c("Closed by arrest","Closed without arrest"), 1, 0)) %>%
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(victim_age = ifelse(victim_age == "unknown", NA, victim_age)) %>%
  mutate(victim_age = as.numeric(victim_age))


cleaned_homicide
```

##### For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.





