---
title: "p8105_hw6_NL2836"
author: "Ngan Le"
date: "2023-12-01"
output: github_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Question 1

##### Import dataset from CSV file. 

```{r import homicide data, message = FALSE}
homicide = read_csv("homicide-data.csv") %>% 
  janitor::clean_names()
```

_This dataset includes `r nrow(homicide)` observations and `r ncol(homicide)` variables, reporting the date, the city, the state, the coordinates of the incidences, the victims' name, age, race, gender, and the disposition of the case. The incidences are reported across `r n_distinct(homicide$state)` states from 2007-2017._

##### Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r create city_state & binary homicide vars, warning = FALSE}
cleaned_homicide =
  homicide %>%
  mutate(city_state = str_c(city, ", ", state)) %>% 
  mutate(solved = ifelse(disposition %in% c("Closed by arrest","Closed without arrest"), 1, 0)) %>%
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(victim_age = ifelse(victim_age == "unknown", NA, victim_age)) %>%
  mutate(victim_age = as.numeric(victim_age))

cleaned_homicide
```

##### For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.


```{r baltimore glm}
baltimore_glm =
  cleaned_homicide %>%
  filter(city_state == "Baltimore, MD") %>%
  glm(solved ~ victim_age + victim_sex + victim_race, 
             data =., 
             family = binomial) %>% 
  broom::tidy(baltimore_glm, 
              exponentiate = TRUE, 
              conf.int = TRUE, 
              conf.level = 0.95) %>% 
  rename(OR = estimate, 
         CI_lower = conf.low, 
         CI_upper = conf.high,
         p_value = p.value) %>%
  mutate(p_value = ifelse(p_value < 0.05, "<0.05", as.character(round(p_value, 3)))) %>%
  select(term, OR, CI_lower, CI_upper, p_value) %>% 
  knitr::kable(digits = 3)

baltimore_glm
```

_With the `female` being the reference category for sex and `Black` being the reference category for race: the odds of having the homicide case solved among male victims is 0.355 times the odds of having homicide case solved among female victims. Since the p-value is smaller than 0.05, we have sufficient evidence to claim that the odds of having the homicide case solved are significantly different between male and female victims._

```{r cities_glm, warning = FALSE}
cities_glm <- cleaned_homicide %>% 
  group_by(city_state) %>%
  nest() %>%
  mutate(
    regression = map(data, ~glm(formula = solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())),
    solved_cities = map(regression, ~broom::tidy(.x, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95))
  ) %>% 
  select(-data, -regression) %>% 
  unnest(solved_cities) %>% 
  filter(term == "victim_sexMale") %>% 
  rename(OR = estimate, 
         CI_lower = conf.low, 
         CI_upper = conf.high,
         p_value = p.value) %>% 
  select(city_state, OR, CI_lower, CI_upper) 

cities_glm
```
# MISSING OR DESCRIPTION

##### Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.


```{r OR plot, dpi = 300}
OR_plot = 
  ggplot(cities_glm, aes(x = reorder(city_state, OR), y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  coord_flip() +
  labs(x = "City", y = "Odds Ratio (OR)", 
       title = "Estimated ORs and CIs for Solving Homicides in Each City") +
  theme_minimal()

OR_plot
```

# MISSING PLOT DESCRIPTION

# Question 2

###### Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂2 and log(β̂0∗β̂1). Note: broom::glance() is helpful for extracting r̂2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂1∗β̂2).

```{r download dataset}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r bootstrap}
model <- lm(tmax ~ tmin + prcp, data = weather_df)

set.seed(123456)

boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

straps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weather_df))
  )
```


```{r bootstrap betas, warning = FALSE}
bootstrap_betas = 
  straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin + prcp, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

log_betas <-  
  bootstrap_betas %>%
  group_by(strap_number) %>%
  summarise(log_betas = log(estimate[2] * estimate[3])) %>%
  select(log_betas, strap_number)
```

```{r bootstrap r2}
bootstrap_r2 <- 
  straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin + prcp, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

r_squared <- 
  bootstrap_r2 %>%
  select(r.squared, strap_number)
```

```{r betas plot, warning = FALSE}
ggplot(log_betas, aes(x = log_betas)) + 
  geom_density() +
  labs(title = "Distribution of log(Beta1 * Beta2)") +
  theme_minimal()
```


```{r r2 plot}
ggplot(r_squared, aes(x = r.squared)) + 
  geom_density() +
  labs(title = "R-squared Distribution") +
  theme_minimal()
```

Mean and SD. 

```{r}
r_squared1 = r_squared %>%
  summarise(r_squared_sd = sd(r.squared), 
            r_squared_mean = mean(r.squared)) %>%
  pull(r_squared_sd, r_squared_mean)

log_betas1 = log_betas %>% 
  summarise(log_betas_sd = sd(as.numeric(log_betas),na.rm = TRUE),
           log_betas_mean = mean(as.numeric(log_betas), na.rm = TRUE) ) %>%
  pull(log_betas_sd, log_betas_mean) 

CI_result <- log_betas %>%
  summarize(ci_lower = quantile(log_betas, 0.025, na.rm = TRUE),
            ci_upper = quantile(log_betas, 0.975, na.rm = TRUE))
```

# MISSING MEAN AND SD, 95% CI DESCRIPTION

